---
title: "Seurat Clustering tutorial"
author: "Poromendro Nath Burman"
date: "1/25/2021"
output: html_document
---

**1. Import libraries**
```{r}
library(dplyr)
library(Seurat)
library(patchwork)
```

**2. Import dataset**
```{r}
# load the dataset
pbmc.data <- Read10X(data.dir = "./filtered_gene_bc_matrices/hg19/")

# create seurat object
pbmc <- CreateSeuratObject(counts = pbmc.data, project = "pbmc3k", min.cells = 3, min.features = 200)

# remove pbmc.data
rm(pbmc.data)
```

**3. Mitochondrial genes**
```{r}
pbmc[["percent.mt"]] <- PercentageFeatureSet(pbmc, pattern = "MT-")
```

**4. QC visualization**

**a) UMI counts**
```{r}
VlnPlot(pbmc, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```

**b) Feature-feature relationships**
```{r}
FeatureScatter(pbmc, feature1 = "nCount_RNA", feature2 = "percent.mt") + FeatureScatter(pbmc, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
```
**5. Filtering of cells**

# Based upon the Violin and feature scatter plots, we filter for cells that have 
# feature counts over 2,500 or less than 200 and >5% mitochondrial counts

```{r}
pbmc <- subset(pbmc, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < 5)
```

**6. Normalization**
```{r}
pbmc <- NormalizeData(pbmc, normalization.method = "LogNormalize", scale.factor = 10000)
```

**7. Feature selection**
```{r}
pbmc <- FindVariableFeatures(pbmc, selection.method = "vst", nfeatures = 2000)

# identify top 10 most highly variable genes
top10 <- head(VariableFeatures(pbmc), 10)

# plot variable features
VariableFeaturePlot(pbmc)
```

**8. Scaling the data**
```{r}
all.genes <- rownames(pbmc)

pbmc <- ScaleData(pbmc, features = all.genes)
```

**9. Dimensional Reduction**

**a) Linear**

# calculate PCA
```{r}
pbmc <- RunPCA(pbmc, features = VariableFeatures(object = pbmc))
```

# PCA visualization
```{r}
print(pbmc[["pca"]], dims = 1:5, nfeatures = 5)
```

```{r}
VizDimLoadings(pbmc, dims = 1:2, reduction = "pca")
```
```{r}
DimPlot(pbmc, reduction = "pca")
```
# heatmap
```{r}
DimHeatmap(pbmc, dims = 1:5, cells = 500, balanced = TRUE)
```
# determining dimensionality
```{r}
ElbowPlot(pbmc)
```
# Clustering
```{r}
pbmc <- FindNeighbors(pbmc, dims = 1:10)
pbmc <- FindClusters(pbmc, resolution = 0.5)
```

**b) Non-Linear**

```{r}
pbmc <- RunUMAP(pbmc, dims = 1:10)
DimPlot(pbmc, reduction = "umap")
```

**c) differentially expressed genes**

# markers of cluster 1
```{r}
cluster1.markers <- FindMarkers(pbmc, ident.1 = 1, min.pct = 0.25)
head(cluster1.markers, n = 5)
```
# markers distinguishing cluster 5 from clusters 0 and 3
```{r}
cluster5.markers <- FindMarkers(pbmc, ident.1 = 5, ident.2 = c(0, 3), min.pct = 0.25)
head(cluster5.markers, n = 5)
```

# find markers for every cluster compared to remaining cells
```{r}
pbmc.markers <- FindAllMarkers(pbmc, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)

pbmc.markers %>% group_by(cluster) %>% top_n(n = 2, wt = avg_logFC)
```

# Plot top 20 markers
```{r}
top10 <- pbmc.markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_logFC)
DoHeatmap(pbmc, features = top10$gene) + NoLegend()
```




